CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.

### 2020-12-10 codecheck-paper.tex
turn notes into content

\begin{itemize}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\item Use "execution" insteand of "sharing" in title?
\item Can we introduce "research compendium" as a term?
\item Should we make code and data sharing a prerequisite or leave that door open?
\end{itemize}

@@

So, assuming that researchers begin to share more artifacts, how might
they be examined to test that they do what they claim? For example,
most scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share.
The implementation status of this varies considerably according to the
journal. At one of the spectrum, there are specialist journals that
have been created to accept ``data papers'' (e.g.~\href{}{\emph{Scientific
  Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}}).
These journals have established rigorous procedures by which
the data is validated according to standards in each field. At the
other end of the spectrum, we still witness today the infamous
statement ``Data available upon reasonable request'' whereby authors,
while possible well-intentioned at the time of writing the article,
often cannot provide the data when readers ask for
it.  (ANDREW H: any references to sum up poor/evolving state of data
sharing; Vines et al 2014).

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? While both data and software
together are required to validate outcomes of a computational analysis,
the do differ in the respect that the former can be seen as static/inert,
while the latter requires an environment and interaction. This makes 
software harder to share.
Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
code cannot be shared, e.g.~``there is no documentation / I do not
want to maintain it / I do not want to give away my code to my
competitors''. Our view, rooted in Open Science princples, is that
wherever possible sharing code is good for the community, as outlined
ten years ago \cite{Barnes2010-iv}. Having the code freely available, and
suitably archived, provides a valuable resource for others to learn
from, even when it doesn't run or if there is no documentation. However,
with a little effort, we believe that if someone independent can run/execute
the code, this is worth documenting as early as possible. With this in
mind, we have developed a set of PRINCIPLES and an example workflow that
provides what he hope is a pragmatic way of checking that code works.

@@

\textbf{Step 1:} the AUTHOR submit their manuscript along with code and data to
the PUBLISHER. The code and data need not be openly available at this point.
However, in many cases the code and data may be published on a code hosting platform,
such as GitHub or GitLab.

@@

\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
the author, and checks if some/all of the results from the paper can be
reproduced. (See later for exact nature of ``reproduced''). If there are
any problems running the code, the CODECHECKER asks the author for help
to resolve the problems. The CODECHECKER then tries to run the code
again. This process iterates until either the CODECHECKER is successful,
or the CODECHECKER concludes the code does not work (\emph{what do in this
case}). As part of this process, the CODECHECKER could work entirely
locally on their own compute resource, or in the cloud, e.g.~using the open
MyBinder infrastructure \cite{jupyter_binder_2018}. Using such cloud-based 
infrastructure allows for the CODECHECKER and author to collaboratively improve
the code and enforces a much more complete definition of the computing environment,
but, unless infrastructure is provided by, e.g., by the publisher, requires code
and data to be published openly online.

\textbf{Step 4:} the CODECHECKER writes a certificate stating how the code was
run and includes a copy of outputs (figures, tables) that were
independently generated. The certificate may include further recommendations to
the author how to improve the material.

@@

\textbf{Step 6:} the PUBLISHER can then use the certificate as part of their
supplementary records for a paper or provide it to scientific reviewers; the
PUBLISHER also gives credit to the CODECHECKER for their work by depositing the
activity appropriately in scholarly publication databases, such as ORCID.
The PULISHER also ensures proper publication metadata, e.g., links from the 
certificate repository to the published paper or the original code repository.

@@

The concrete people to provide these expertises can range from researchers,
possibly drawn from the regular pool of volunteer reviewers or from a 
special group, to editorial staff of a publisher. If the regular reviewers
are relied upon, the editor's task to find suitable ones becomes more
tedious; extra information such as software familiarity would have to be
modelled in review platforms.
We see a great opportunity to especially involve early career 
researchers as codecheckers. They arguably have a high interest in learning
about the latest tools and technologies as they are building up their own
expertise and specialisation. They are also introduced to peer review and 
may transition into the role of scientific reviewer over time.
Not to solve the general problems of peer review, which is largely an
unsupervised process, i.e., rarely are reviewers taught how to do it, but
a junior codechecker may be supported by a senior codechecker, too.
We see a high potential in setting up the new processes for CODECHECKING with
a clear committment on opennes and transparency, independent of the
respective current peer review process (see \nameref{who-knows-who}).

@@

In contrast, for researchers it can be very important to be publicly credited
with their activity as a reviewer (see \nameref{who-knows-who}). A regular
review may be listed in public databases (e.g., ORCID
\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
\footnote{\href{https://publons.com/}{https://publons.com/}} or 
ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
A codechecker should be listed just the same. The number of scientific reviews
if of course a very coarse indicator for the community contribution by an 
individual researcher, so it's unclear if a regular reviewer, who in addition
conducts a CODECHECK to the scientific review, should be credited twice.

@@

As a prerequisite, we see a CODECHECK as part of a typical scholarly
communication process. While more radical ideas to change the paradigms
how researchers share their work exist\footnote{For example Octopus, 
\url{https://science-octopus.org/about}, or Hypergraph
\url{https://www.libscie.org/hypergraph}.}, there is a 
value in an evolutionary approach. Thereby, existing communities can 
transition towards more open practices at their own pace and ensure
no one is left behind unfairly.
When integrating a CODECHECK in existing review and publication processes,
the \emph{turnaround time} is crucial. Depending on when and who conducts
the check, it may be done faster, or add considerably to the overall
time from submission to publication.
% TODO write about our experiences how long a CODECHECK takes?

@@

\textbf{4. Workflows must be auditable.} This requires that the codechecker 
has at least enough material to validate the workflow outputs submitted by 
the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'.
Code could be executed at least once without critical errors or
warnings using the provided instructions, and, therefore, all code and data 
was given and could be investigated more deeply or extended in the future.
Ideally, this is a "one~click" step, but achieving this requires particular 
skills. It is also hard to create a sufficient level of documentation for 
third parties. That is why the possibility to communicate is more important
than perfected workflows, and why a CODECHECK is not automated.
A codechecker does \emph{not} conduct a code review.
% more content from the website, may fit better elsewhere:
% The CODECHECK is not automated on purpose: automation may (a) lead to people gaming the system, (b) hide details that eventually decrease level of certainty that a codechecker has in their assessment, and (c) reduce the understandability of instructions in the long term, which is more important than short term ease of use (see Principle 2).

\textbf{5. Open by default.}  By default, unless there are strong
reasons to the contrary (e.g., clinically-sensitive data), all code and
data, both from author and codechecker, will be made freely available, i.e., 
under open licenses, before or when the certificate is published.
The code and data  publication should follow community good practices.

@@

We have a curated list of 22 certificates available at
\url{https://codecheck.org.uk/register}. These are a mixture of
reproductions of papers for a variety of reasons and fall into three themes:
(1) `classic' papers from computational neuroscience
(2) COVID-19 modelling preprints, and
(3) GIScience.  
The first theme was the initial set of papers used to explore the concept
of CODECHECK. The idea was to take very well known historical articles from a 
specific domain, in this case the work area of co-author S.~Eglen, possibly 
as as a conversation starter for potential collaborating journals.
The second theme is rather opportunistic. As the COVID-19 pandemic swept
over the whole Earth, we hoped to contribute a small part to science's answer
to the unknown virus. The checks were solicited through community interaction
%Several COVID-19 models have been done during the time of peer review of these papers,
and under our initative rather than requested from journal, but the certificate
¸has then been acknowledged/cited in the paper (LANCET/LSHTM). % TODO SJE: clarify references in brackets
The third theme represents co-author D.~N\"ust's service as Reproducibility
Reviewer at the AGILE conference series.
Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
established a process for reproducing workflows at the AGILE conference
series \cite{nust_improving_2020}.
While using slightly different terms and infrastructue, e.g., ``reproducibility
reports'' are published on OSF instead of certificates on Zenodo, the 
reproducibility reviews adhere to the codecheck principles.
Furthermore, a small number of checks was completed as part of a peer reviews
for a GIScience journals.

@@

These tools and resources demonstrate that a CODECHECK process can be
managed without infrastructure costs on freely available platforms.
The scripted and automated parts should also ensure that more checks can
be conducted in the future.
The remaining costs are the project's domain, which is negligible, and
the human resources for managing the project and processes, which are
considerable and currently soley based on, partly grant-based, public funding.
Naturally, similar set-ups are possible, such as using GitLab/GitLab.com 
and the GitLab CI instead of GitHub and GitHub Actions, or using a different 
data repository for depositing certificates and data.

@@

TO BE CONTINUED.

All the related projectsfrom A-Z\ldots{} e.g.~CASCAID, Rescience,
modeldb, o2r. Community iniatives like reprohack along similar lines of
checking code. \url{https://reprohack.github.io/reprohack-hq/}

\url{https://twitter.com/Code_Copilot}

% first glimpse into "Code exeuction processes in scholarly publishing" ?
% `https://docs.google.com/document/d/1fMWrFvBTwAYXg5J_bYIjgb-gY7DfQ0DOjx9Vlg_Rbmg/edit# ?

DELIBERATEY NOT COMPREHENSIVE HERE.

@@

\textbf{Not going deep enough}: 
It should be noted that a codechecker is not taking the same roles as the
\emph{statistical reviewer}, as it is practices by some journals in the 
biomedical domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).
The statistical reviewer actually evaluates the appropriateness of
statistical methods \cite{greenwood_how_2015} and can support topical
reviewers if, e.g., complex methods or sophisticated variants of statistical
tests are applied \cite{petrovecki_role_2009}.
The codechecker may go equally deep into the review, if they have, by 
intention of by chance, the required expertise and the required time, 
but they are by no means expected to be able to provide
such a level of scrutiny. We can imagine a tiered workflow where the 
codechecker, just as the conventional reviewer, could recommend a detailed
statistical review to the editor as they might come across upon different
hints.
%The CODECHECK provides a certain level of safety with regard to the inherent
%problems that a computational workflow might have. We want to capture the
%obvious errors, but ... https://www.quora.com/What-is-the-difference-between-safety-and-security

\textbf{Handling failure}: we have no established process for the case
that a reproduction fails. This may be the sign of a pre-selection bias
of the examples, as so far all our codechecks succeeded.
In case of a journal adopting CODECHECK for all submissions, the open 
question remains as what to do in this case will have to be answered.
At the start, we doubt publicly reporting failures (i.e., the code 
would not run) will increase overall reproducibility and recommend not
to share the outcome publicly but to share the latest draft of the 
certific with the authors.
One must be careful as to the impact of acceptance by authors and the
concerns of volunteers this approach, or a contrary one, has.

\textbf{Compute time}: for those papers that take significant compute
time (think days, not minutes), who will pay for the compute time?
The openness of CODECHECK certificates could help to establish community
peer pressure, as submitting authors can be evaluated on their engagement
as codecheckers.

\textbf{Proprietary software}: authors can currently provide code that requires
proprietary software. Given the prevalence of proprietary software in
some disciplines, such as MATLAB or eCognition,
this seems like a pragmatic choice.
However, it prohibits to benefit from open infrastructure for reproducibility
(cf.~\cite{konkol_publishing_2020}) and requires the codechecker to have access 
to that software.
Non-open software also considerably hampers reuse, especially by researchers
from the global south. Therefore, allowing proprietary software is a compromise
that each implementation of CODECHECK should critically reconsider.

>>>>

\begin{itemize}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\item Use "execution" insteand of "sharing" in title?
\item Can we introduce "research compendium" as a term?
\item Should we make code and data sharing a prerequisite or leave that door open?
\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
\item Should we ask the people currently listed in the acknowledgements?
\item revisit \url{https://github.com/codecheckers/discussion}
\end{itemize}

@@

So, assuming that researchers begin to share more artifacts, how might
they be examined to test that they do what they claim? For example,
most scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share.
The implementation status of this varies considerably according to the
journal. At one of the spectrum, there are specialist journals that
have been created to accept ``data papers''
(e.g.~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}}).
These journals have established rigorous procedures by which
the data is validated according to standards in each field. At the
other end of the spectrum, we still witness today the infamous
statement ``Data available upon reasonable request'' whereby authors,
while possible well-intentioned at the time of writing the article,
often cannot provide the data when readers ask for
it.  (ANDREW H: any references to sum up poor/evolving state of data
sharing; Vines et al 2014).

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? While both data and software
together are required to validate outcomes of a computational analysis,
the do differ in the respect that the former can be seen as static/inert,
while the latter requires an environment and interaction. This makes 
software harder to share.
Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
code cannot be shared, e.g.~``there is no documentation / I do not
want to maintain it / I do not want to give away my code to my
competitors''. Our view, rooted in Open Science princples, is that
wherever possible sharing code is good for the community, as outlined
ten years ago \cite{Barnes2010-iv}. Having the code freely available, and
suitably archived, provides a valuable resource for others to learn
from, even when it doesn't run or if there is no documentation. However,
with a little effort, we believe that if someone independent can run/execute
the code, this is worth documenting as early as possible and that the barrier
to evaluating non-text parts of research can be greatly reduced with a 
focused approach.
Just like peer review can provide a ``baseline reasurance'', i.e., a paper
has at least ben checked by someone with an understanding of the topic
\cite{fyfe_mission_2019}, we think the same baseline should be provided
regarding the workflow underlying a paper.
With this in mind, we have developed a set of PRINCIPLES and an example 
workflow that provides what he hope is a pragmatic way of checking that 
code works.

@@

\textbf{Step 1:} the AUTHOR submit their manuscript along with code and data to
the PUBLISHER. The code and data need not be openly available at this point.
However, in many cases the code and data may be published on a code hosting platform,
such as GitHub or GitLab. Ideally, the AUTHOR is expecting the CODECHECK and
prepares for it, e.g., by asking a colleague to briefly attempt a reproduction
of the submitted workflow.

@@

\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
the author, and checks if some/all of the results from the paper can be
reproduced. (See later for exact nature of ``reproduced''). If there are
any problems running the code, the CODECHECKER asks the author for help
to resolve the problems. The burden to provide reproducible material lies with
the author.
The CODECHECKER then tries to run the code again.
This process iterates until either the CODECHECKER is successful,
or the CODECHECKER concludes the code does not work (\emph{what do in this
case}). As part of this process, the CODECHECKER could work entirely
locally on their own compute resource, or in the cloud, e.g.~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which
are more tailored for scientific publications while others offer commercial
options to, e.g., publishers \cite{}.
Using such cloud-based 
infrastructure allows for the CODECHECKER and author to collaboratively improve
the code and enforces a much more complete definition of the computing environment,
but, unless infrastructure is provided by, e.g., by the publisher, requires code
and data to be published openly online.
Note that the taks of the CODECHECKER is to check the ``mechanics'' of the
workflow, not to evaluate its correctness. 
Stodden et~al. \cite{stodden_setting_2013} destinguish between verification and
validation in the context of Mathematics, and in the same vein a CODECHECK
ensures a verification\footnote{Note this is a different use as the \emph{Verification Reports} article type in the journal \emph{Cortex}
\cite{chambers_verification_2020}},
i.e., checking that the code correctly produces the 
output it claims to create, but not a validation, i.e., does the code implement
the right algorithm to solve the problem that is researched, 
of the computational results.

\textbf{Step 4:} the CODECHECKER writes a certificate stating how the code was
run and includes a copy of outputs (figures, tables) that were
independently generated. The certificate may include further recommendations to
the author how to improve the material.
The plain text in the certificate is the most flexible and feasible way to 
make clear what was really checked. Readers will have very different 
understandings of what a ``green'' checkmark (cf. Step~5) means.

@@

\textbf{Step 6:} the PUBLISHER can then use the certificate as part of their
supplementary records for a paper or provide it to scientific reviewers; the
PUBLISHER also gives credit to the CODECHECKER for their work by depositing the
activity appropriately in scholarly publication databases, such as ORCID.
The PULISHER also ensures proper publication metadata, e.g., links from the 
certificate repository to the published paper or the original code repository.
A badge or other visual aid may be applied to highlight the fact of a completed
CODECHECK and provide a incentive and reward for the, today often considered
``additional'' efforts of providing readily computationally reproducible workflows,
despite the simplification into a binary value and the possible confusion of
the meaning, especially across disciplines and methods.

@@

The concrete people to provide these expertises can range from researchers,
possibly drawn from the regular pool of volunteer reviewers or from a 
special group, to editorial staff of a publisher. If the regular reviewers
are relied upon, the editor's task to find suitable ones becomes more
tedious; extra information such as software familiarity would have to be
modelled in review platforms. Keeping diverse specialists as staff to
realise a codechking process is unlikely to be feasible, especially for 
small or independent publishers.

We consider a single codechecker to be enough, unlike peer review where
multiple referees are common. Because of the technical focus, the 
determination if the concrete outputs of author and codechecker, who can
collaborate to solve any issues, should be a very factual process and not
require much debate or opinion. Depending on the power of the CODECHECK
(see \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
second opinion might be warranted in rare cases.

We see a great opportunity to especially involve early-career researchers
(ECRs) as codecheckers. They arguably have a high interest in learning
about the latest tools and technologies as they are building up their own
expertise and specialisation. The possibility to gain insights into latest
science for ECRs and at the same time increase the appreciation of
reproduction efforts is a sweet spot that CODECHECK shares
with \emph{ReScience X}, a journal devoted to reproduction and replication 
experiments \cite{roesch_new_2020}.
In fact, ECRs are often much more familiar with latest technology and are 
likely to be early adoptors as creators, i.e., be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
submitted to leading Neuroscience journals by ECRs
\cite{chambers_registered_2019}.}
They are also introduced to peer review and 
may transition into the role of scientific reviewer over time.
Not to solve the general problems of peer review, which is largely an
unsupervised process, i.e., rarely are reviewers taught how to do it, but
a junior codechecker may be supported by a senior codechecker, too.
We see a high potential in setting up the new processes for CODECHECKING with
a clear committment on openness and transparency, independent of the
respective current peer review process (see \nameref{who-knows-who}).

@@

In contrast, for researchers it can be very important to be publicly credited
with their activity as a reviewer (see \nameref{who-knows-who}). A regular
review may be listed in public databases (e.g., ORCID
\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
\footnote{\href{https://publons.com/}{https://publons.com/}} or 
ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
A codechecker should be listed just the same. The number of scientific reviews
is of course a very coarse indicator for the community contribution by an 
individual, yet is is unclear if a regular reviewer, who in addition
conducts a CODECHECK to the scientific review, should be credited twice.

@@

As a prerequisite, we see a CODECHECK as part of a typical scholarly
communication process. While more radical ideas to change the paradigms
how researchers share their work exist\footnote{For example Octopus, 
\url{https://science-octopus.org/about}, or Hypergraph
\url{https://www.libscie.org/hypergraph}.}, there is a 
value in an evolutionary approach. Thereby, existing communities can 
transition towards more open practices at their own pace and ensure
no one is left behind unfairly.
When integrating a CODECHECK in existing review and publication processes,
the \emph{turnaround time} is crucial. Depending on when and who conducts
the check, it may be done faster, or add considerably to the overall
time from submission to publication.
In our first checks, we found that a codecheck may generally take from 2-5
hours, with some outliers on the higher end. This time includes writing and
publishing the certificate, but excludes the time for actual computations.
The efforts include considerable time to communicate about the process, 
especially around who publishes what and when, so that proper 
cross-referencing between paper and certificate is ensured.
When integrated into peer review systems, this handling of the steps should
become much more effective.

@@

\textbf{4. Workflows must be auditable.}
This requires that the codechecker 
has at least enough material to validate the workflow outputs submitted by 
the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'
and the ICERM report \cite{} defines the level ``Auditable Research'' 
similarly. Note that openness is not required here, as to accommodate 
communities of practice and review workflows that (currently) do not want
to or are not able to work openly.
A completed check eans that code could be executed at least once without
critical errors or
warnings using the provided instructions, and, therefore, all code and data 
was given and could be investigated more deeply or extended in the future.
Ideally, this is a "one~click" step, but achieving this requires particular 
skills. It is also hard to create a sufficient level of documentation for 
third parties. Furthermore automation may lead to people gaming the system
or reliance on technology hiding important details. All such aspects can
reduce the understandability of the material, which we estimate to be the
best mean to ensure long-term transparency and usefulness.
That is why the possibility to communicate is more important
than perfected workflows, and why a CODECHECK is not automated.
We acknowledge that others have argued in favour of bitwise reproducibility
because in the long run it can be automated\footnote{Konrad Hinsen on 
Twitter: \url{https://twitter.com/khinsen/status/1242842759733665799}};
but until then we need CODECHECK's approach.

\textbf{5. Open by default and transitional by disposition.} % or "design"?
By default, unless there are strong
reasons to the contrary (e.g., clinically-sensitive data), all code and
data, both from author and codechecker, will be made freely available, i.e., 
under open licenses, before or when the certificate is published.
The code and data  publication should follow community good practices.
By disposition, a CODECHECK workflow should be designed as a recognition of
shortcomings in asserting reproducibility of submissions and understand the
complex situation that lead to this. However, establishing a CODECHECK workflow
should go hand in hand with initiatives and a perspective to improve education
around reproducibility of methods and communication of comprehensive 
computational workflows, with the long term goal for a specific community or
journal's target audience to reach a state where the level of verification 
and trust provided by codecheckers becomes part of peer review.

@@

We have a curated list of 22 certificates available at
\url{https://codecheck.org.uk/register}. These are a mixture of
reproductions of papers for a variety of reasons and fall into three themes:
(1) `classic' papers from computational neuroscience
(2) COVID-19 modelling preprints, and
(3) GIScience.  
The first theme was the initial set of papers used to explore the concept
of CODECHECK. The idea was to take very well known historical articles from a 
specific domain, in this case the work area of co-author S.~Eglen, possibly 
as as a conversation starter for potential collaborating journals.
The second theme is rather opportunistic. As the COVID-19 pandemic swept
over the whole Earth, we hoped to contribute a small part to science's answer
to the unknown virus. The checks were solicited through community interaction
%Several COVID-19 models have been done during the time of peer review of these papers,
and under our initative rather than requested from journal, but the certificate
has then been acknowledged/cited in the paper (LANCET/LSHTM). % TODO SJE: clarify references in brackets
The third theme represents co-author D.~N\"ust's service as Reproducibility
Reviewer at the AGILE conference series.
Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
established a process for reproducing workflows at the AGILE conference
series \cite{nust_improving_2020}.
While using slightly different terms and infrastructue, e.g., ``reproducibility
reports'' are published on OSF instead of certificates on Zenodo, the 
reproducibility reviews adhere to the codecheck principles.
Furthermore, a small number of checks was completed as part of a peer reviews
for a GIScience journals.

@@

These tools and resources demonstrate that a CODECHECK process can be
managed without infrastructure costs on freely available platforms.
The scripted and automated parts should also ensure that more checks can
be conducted in the future, though the management by a paid CODECHECK
editor or further automation, e.g., with a bot, could improve turnaround time.
The remaining costs are the project's domain, which is negligible, and
the human resources for managing the project and processes, which are
considerable and currently soley based on, partly grant-based, public funding.
Naturally, similar set-ups are possible, such as using GitLab/GitLab.com 
and the GitLab CI instead of GitHub and GitHub Actions, or using a different 
data repository for depositing certificates and data.

@@

The journal \emph{ACM Transactions on Mathematical Software (TOMS)} 
established a "Replicated Computational Results" (RCR) review proces
\cite{heroux_editorial_2015} (using "replicable" with the same meaning 
as CODECHECK uses "reproducible").
The CODECHECK principles were independently developed as we only later 
learned about TOMS RCR, which we judge as fulfilling CODECHECK principles
1 to 4.
This work provides more extensive discussion of the possible variants, 
while TOMS RCR realises one particular workflow for reproducibility 
reviews for one particular journal. \cite{heroux_editorial_2015}
include similar concerns about
selection of reviewers as above. Unlike existing CODECHECK certificates, 
the RCR reports undergo editorial review, but this could very well become
part of CODECHECK workflows at journals.
The publication of the RCR report is seen as a reward for the efforts of
the reproducing person, while the possible conflict of interest of this 
motivation is acknowledged. TOMS also reconises reviewer activity in a 
partnership with Publons
\footnote{\url{https://authors.acm.org/author-services/publons}},
though it is unclear if the author of an RCR report receives both the report
authorship and the reviewer recognition. In any case, this is different
from CODECHECK, which specifically aims at recognition as reviewers.
In our view, this removes the possible conflict of interest while keeping
the public acknowledgement.

Specific to the field of Mathematics, the RCR was also  expected to apply a
review of the software itself if the system it runs on cannot be evaluated by
an independent party.
The TOMS RCR also notes the importance of communication and expects
collaboration between author and RCR reviewers.
The TOMS creators share thoughts around reviewer selection and also put trust
in reviewer judgement over numerical bit-wise perfection.
A key difference is that authors opt-in with an \emph{RCR Review Request} and
the RCR reports, which are very similar to CODECHECK certificates, are
published in the TOMS journal next to the actual papers.
A search on Crossref\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2}; search executed
on 2020-12-10} shows 15 RCR Reports to be published and the workflow to be 
extended to the ACM journal \emph{Transactions on Modeling and Computer 
Simulation (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines#rcrini}}.

% 8: https://scholar.google.de/scholar?start=10&q=Replicated+Computations+Results+%22(RCR)+Report%22&hl=de&as_sdt=0,5
% 15: https://search.crossref.org/?q=Replicated+Computations+Results+%22%28RCR%29+Report%22&from_ui=yes&publication=ACM+Transactions+on+Mathematical+Software
% 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10

A number of journals or conferences/workshops provide guidelines for ensuring reproducible publications and their verification.
These include, for example, \ldots{}
% first glimpse into "Code exeuction processes in scholarly publishing" ?
% `https://docs.google.com/document/d/1fMWrFvBTwAYXg5J_bYIjgb-gY7DfQ0DOjx9Vlg_Rbmg/edit# ?

\emph{Information Systems}, with an invited reproducibility paper track
\footnote{\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}}

However, no approach besides CODECHECK spans across journals and is
intentionally designed to be picked up and implemented by many different 
journals or events, and to build a community of codecheckers. CODECHECK shares
this interdisciplinary nature with other community iniatives concered with 
reproducibility awareness, education, and supports, such as ReproHack
\footnote{\url{https://reprohack.github.io/reprohack-hq/}}, Code Copilot
\footnote{\url{https://twitter.com/Code_Copilot}}, or Papers with Code
\footenote{\url{https://paperswithcode.com/about}}.
The latter recently announced a collaborations with the preprint server
\emph{arXiv} on providing data and code supplements for machine learning
manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}} and runs a reproduciblity challenge
\footnote{\url{https://paperswithcode.com/rc2020}}.

The journal \emph{ReScience C} publishes replications,
\url{https://rescience.github.io/faq/};
they do not accept proprietary software and require open source code;
they also required a third party to replicate; they also use Zenodo
to issue DOIs.
The journal \emph{Cortex} has a special article type 
\emph{Verification Reports}, which  actually are about replication of results 
but are very well designed/reasoned \cite{chambers_verification_2020}.
In a similar vein, the CODECHECK certificates
could also be published as a special article type within journals.

For research with very high stakes, where reproduction would be too weak and
post publication replication possibly to late because of policiy impact,
\cite{benjamin-chung_internal_2020} propose \emph{internal replication}.
A workflow that underwent internal
replication would very likely be of high quality and relatively easy to codecheck.
In a similar vein, internal codechecks may be used, with the same limitations such
as group think \cite{benjamin-chung_internal_2020},
to ensure reproducibility before submission. The guidelines
for CODECHECK can be used by fellow/not independent lab members just as well.

Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline and 
infrastructure for reproducible computational research with its own tools and
API. Their specific packaging format, provenance record, and cryptographic
\emph{Verifiable Result Identifier} would indeed provide excellent
reproduciblity. However, the system is also complex and was not taken up by
any publisher as far as we know; the system is indeed not open source itself.
In comparison, CODECHECK is less powerful but also much more flexible and
less dependent on specific tools or infrastructure. If data and code are 
deposited properly, i.e., very unlike to disappear, then the certificate's 
DOI is practically quite close to the cryptographic identifier.

Local reproduction initiatives/services:
\href{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/}{CISER R-squared}
(established/permanent),
\href{https://go.wwu.de/r2s2}{WWU R2S2} (pilot only, focused on reproductions),
\href{https://github.com/OxfordCodeReviewNet/forum}{Oxford code review network (OxCRN)}
(community, broader scope).
For these, the CODECHECK scope, tools, and practical experiences may be
helpful resources.

\emph{SciGen.Report} is a community-run platform to foster communication on
reproducibility\footnote{\url{https://scigen.report/}}. People can report on
fully, particially or un-successful reproductions post publication.

@@

\textbf{Not going deep enough/too vague}:
CODECHECK by design does not require authors to provide (and sustain) an
eternally functional worklow and neither suggests a specific software stack
or pracitcal approaches, such as research compendia
\footnote{\url{https://research-compendium.science/}},
even if they are very valuable.
Creating something that anyone can reproduce has been called a 
fool's errand
\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
and we tend to agree.
However, the package of data, code, and documentation
collaboratively created by authors and codecheckers is a snapshot of a 
working analysis that greatly increases the likelyhood of a successful 
reproduction and possibility to extensions by third parties, with access
to suitable resources and matching skill set, in the future.
Concrete implementations of CODECHECK workflows, especially for specific
disciplines, may reify much more helpful guidelines for authors how to
create reproducibility packages.
The author-friendly ``low bar'' should not stay low forever, but cultural
change takes time and the encouragement and guidance that CODECHECK,
as part of the widely accepted peer review concept, can provide may
eventually allow to raise the bar as high as possible, e.g., with
executable research compendia \cite{nust_opening_2017}
or continuous analysis \cite{beaulieu-jones_reproducibility_2017-1}.
However, considering that missing artefacts and lack of documentation
have repeatedly been identified as main barriers to reproducibility
(e.g., \cite{stagge_assessing_2019,nust_improving_2020}),
we would not underestimate the power of a simple check.

It should be noted that a codechecker is not taking the same roles as the
\emph{statistical reviewer}, as it is practices by some journals in the 
biomedical domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).
The statistical reviewer actually evaluates the appropriateness of
statistical methods \cite{greenwood_how_2015} and can support topical
reviewers if, e.g., complex methods or sophisticated variants of statistical
tests are applied \cite{petrovecki_role_2009}.
The codechecker may go equally deep into the review, if they have, by 
intention of by chance, the required expertise and the required time, 
but they are by no means expected to be able to provide
such a level of scrutiny. We can imagine a tiered workflow where the 
codechecker, just as the conventional reviewer, could recommend a detailed
statistical review to the editor as they might come across upon different
hints.
%The CODECHECK provides a certain level of safety with regard to the inherent
%problems that a computational workflow might have. We want to capture the
%obvious errors, but ... https://www.quora.com/What-is-the-difference-between-safety-and-security

An important distinction is that a codechecker does not conduct a 
\emph{code review}. Code reviews are without doubt extremely valuable in 
terms of improving reproducibility and re-usability, and their proponents 
even believe they can improve the science \cite{petre_code_2014}.
Code reviews, however, have quite different structural challenges and 
require even more resources. That said, a well-reviewed codebase is likely
to be easier to codecheck, and the awareness for high-quality code raised
through CODECHECK may lead to more support for code reviewing.
Initiatives and journals that conduct software reviews independently of 
a specific publication or workfloware, e.g.,
ROpenSci\footnote{\url{https://ropensci.org/}},
PyOpenSci\footnote{\url{https://www.pyopensci.org/}},
and JOSS\footnote{\url{https://joss.theoj.org/}}.

Furthermore, the codechecker's task list is intentionally not overloaded
with related concerns, even though they are very important, such as ensuring
proper citation of data and software, or depositing data and software in 
suitablerepositories.
We do expect codecheckers to be aware of these though and possibly point 
them out.
